\documentclass[12pt,a4paper]{article}
\usepackage[margin=2.6cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{booktabs,array}
\usepackage{enumitem}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black,
  pdftitle={Universal Origins: Zero-Sum Constraint},
  pdfauthor={Romeo Matshaba},
  pdfsubject={Theoretical Physics / Cosmogenesis},
  pdfkeywords={zero-sum constraint, cosmogenesis, Planck scale, Standard Model}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{axiom}{Axiom}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{openq}[theorem]{Open Question}

% ── Macros ────────────────────────────────────────────────────────────────────
\newcommand{\EPl}{E_{\mathrm{Pl}}}
\newcommand{\tPl}{t_{\mathrm{Pl}}}
\newcommand{\lPl}{\ell_{\mathrm{Pl}}}
\newcommand{\TPl}{T_{\mathrm{Pl}}}
\newcommand{\MPl}{M_{\mathrm{Pl}}}
\newcommand{\gstar}{g_{*}}
\newcommand{\ZeroMat}{\mathbf{0}_{\infty}}
\newcommand{\rhohat}{\hat{\rho}}
\newcommand{\Qhat}{\hat{Q}}
\newcommand{\Hhat}{\hat{H}}
\newcommand{\Fock}{\mathcal{F}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\ZSC}{\mathcal{Z}}
\newcommand{\TTUO}{T_{\mathrm{TUO}}}

\pagestyle{fancy}
\fancyhf{}
\rhead{\small Universal Origins: Zero-Sum Constraint}
\lhead{\small Matshaba (2026)}
\cfoot{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ══════════════════════════════════════════════════════════════════════════════
\begin{document}
% ══════════════════════════════════════════════════════════════════════════════

\begin{titlepage}
\centering\vspace*{1.8cm}
{\LARGE\bfseries
Universal Origins: The Zero-Sum Constraint\\[0.4em]
Matrix Formulation of Cosmogenesis}\\[1.2em]
{\large\itshape
Infinite-Dimensional Linear Algebra Structure of the Pre-Planck Vacuum\\
and Its Consequences for Big Bang Initial Conditions}\\[2em]
{\large Romeo Matshaba}\\[0.4em]
{\normalsize Department of Physics, University of South Africa, Pretoria, South Africa}\\[0.4em]
{\normalsize 2026}\\[2em]
\rule{0.5\textwidth}{0.4pt}\\[1.5em]
\begin{abstract}
\noindent
We develop a matrix-algebraic framework for cosmogenesis based on a single
structural constraint on the universe's density operator.
Let $\Fock = \bigoplus_{n=0}^{\infty}\Hilb_n$ be the Fock space of possible
universe states, $\rhohat(t) \in \mathcal{B}(\Fock)$ the global density
operator, and $\{\Qhat_k\}$ the complete set of Standard Model conserved
charges.
We impose one axiom: $\Tr[\rhohat(t)\Qhat_k] = 0$ for all $k$ and all times
$t$ --- a constraint that every row of the observable vector
$\mathbf{Q}[\rhohat]$ must equal zero, which we write
$\mathbf{Q}[\rhohat] = \ZeroMat$ (the infinite zero matrix).

Working strictly within this axiom and the Heisenberg uncertainty principle
applied to the Standard Model degree-of-freedom count $\gstar = 106.75$,
we derive --- without additional assumptions --- the following results that
standard Big Bang cosmology takes as inputs:
the radiation equation of state $w = 1/3$,
the Hubble rate $H(\tPl) = 1/(2\tPl)$,
exact spatial flatness $\Omega = 1$,
and the expansion law $V(t) = (4\pi/3)\lPl^3[1+(ct/\lPl)^2]^{3/2}$.
We also present a path-integral saddle-point analysis suggesting that
simultaneous emergence of all Planck cells is exponentially preferred
over localised fluctuations as a possible mechanism for CMB uniformity,
though a complete proof is left as an open problem.
Throughout, we distinguish clearly between results that follow from
the axiom and those that require additional physics beyond TUO's present
scope.
\end{abstract}
\end{titlepage}

\tableofcontents
\newpage
\onehalfspacing

% ══════════════════════════════════════════════════════════════════════════════
\section{Introduction}
% ══════════════════════════════════════════════════════════════════════════════

The two deepest unanswered questions in cosmology are: \emph{why did the
universe begin?} and \emph{why did it begin with those particular initial
conditions?}
The second question is, in a sense, more tractable.
Standard Big Bang cosmology describes the evolution of the universe
extraordinarily well from $t \sim \tPl$ onwards, but its initial
conditions --- the equation of state $w = 1/3$, the spatial flatness
$\Omega = 1$, and the temperature $T \approx \TPl$ --- are inputs rather
than outputs~\cite{Weinberg2008,Kolb1990}.
The inflationary scenario addresses some of these by introducing an
additional scalar field, but it does not eliminate the question of
pre-inflationary initial conditions~\cite{Guth1981,Linde1982}.

The present paper develops the Theory of Universal Origins (TUO),
which proposes that the relevant initial conditions follow mathematically
from a single structural constraint on the universe's density operator
in an infinite-dimensional Fock space.
The constraint is expressed as: the trace of the density operator against
any conserved charge operator must equal zero at all times.
Symbolically, the observable vector built from all conserved charges
equals the zero object --- the infinite zero matrix.
Everything, in a precise operator-algebraic sense, sums to nothing.

The mathematical language we adopt is that of infinite-dimensional linear
algebra.
The density operator is an element of $\mathcal{B}(\Fock)$; the charge
operators are its diagonal entries in an appropriate basis; the constraint
is a collection of linear equations on those diagonal entries.
This language makes the structure of the constraint --- and what it does and
does not determine --- maximally transparent.

We emphasise what this paper does \emph{not} claim:
\begin{itemize}[leftmargin=2em]
  \item We do not claim to explain the number of fermion generations.
  \item We do not claim to derive the baryon asymmetry $\eta$
        quantitatively.
  \item We do not claim to explain the CMB power spectrum.
  \item We do not claim to replace general relativity; a companion
        framework (Quantum Gravity Dynamics) that recovers GR from the
        flat background is required and is stated as a separate project.
  \item We do not derive dark matter or dark energy from these axioms.
  \item Where a result requires physics beyond the two axioms, we
        explicitly label it as a \emph{possible mechanism} or
        \emph{open problem}.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Mathematical Foundations: The Zero Matrix and Fock Space}
\label{sec:math}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{The zero object}

The central object of TUO is the infinite zero matrix:
\begin{equation}\label{eq:zero_matrix}
  \ZeroMat = \begin{pmatrix}
    0 & 0 & 0 & \cdots \\
    0 & 0 & 0 & \cdots \\
    0 & 0 & 0 & \cdots \\
    \vdots & \vdots & \vdots & \ddots
  \end{pmatrix}
  \in \mathcal{B}(\Fock).
\end{equation}
The claim of TUO is that the observable content of the universe, when
correctly assembled, equals $\ZeroMat$ at every instant.
This is not a dynamical statement; it is an existence filter.

\subsection{Fock space structure}

\begin{definition}[Universe Fock space]
Let $\Fock = \bigoplus_{n=0}^{\infty}\Hilb_n$ be the bosonic-fermionic
Fock space over the single-particle Hilbert space $\Hilb_1$ of Standard
Model fields.
States $|\Psi\rangle \in \Fock$ are superpositions of $n$-particle sectors:
\begin{equation}
  |\Psi\rangle
  = \begin{pmatrix} c_0 \\ c_1 \\ c_2 \\ \vdots \end{pmatrix},
  \quad
  \sum_{n=0}^\infty |c_n|^2 = 1.
\end{equation}
\end{definition}

\subsection{Density operator}

The universe is described by a density operator
$\rhohat(t) \in \mathcal{B}(\Fock)$ satisfying
\begin{equation}\label{eq:rho_props}
  \rhohat = \rhohat^\dagger,
  \quad
  \Tr[\rhohat] = 1,
  \quad
  \langle\psi|\rhohat|\psi\rangle \geq 0 \;\forall\,|\psi\rangle.
\end{equation}
In a basis $\{|n\rangle\}_{n=0}^\infty$ of $\Fock$:
\begin{equation}\label{eq:rho_matrix}
  \rhohat = \sum_{m,n=0}^\infty \rho_{mn}|m\rangle\langle n|
  = \begin{pmatrix}
    \rho_{00} & \rho_{01} & \rho_{02} & \cdots \\
    \rho_{10} & \rho_{11} & \rho_{12} & \cdots \\
    \rho_{20} & \rho_{21} & \rho_{22} & \cdots \\
    \vdots & \vdots & \vdots & \ddots
  \end{pmatrix}.
\end{equation}

\subsection{Charge operators}

Each conserved charge corresponds to a Hermitian operator
$\Qhat_k \in \mathcal{B}(\Fock)$ with eigenvalues $\{q_{kn}\}_{n\geq 0}$.
In the eigenbasis of $\Qhat_k$:
\begin{equation}\label{eq:Q_matrix}
  \Qhat_k = \sum_{n=0}^\infty q_{kn}|n\rangle\langle n|
  = \begin{pmatrix}
    q_{k0} & 0 & 0 & \cdots \\
    0 & q_{k1} & 0 & \cdots \\
    0 & 0 & q_{k2} & \cdots \\
    \vdots & \vdots & \vdots & \ddots
  \end{pmatrix}.
\end{equation}
The expectation value is the trace:
\begin{equation}\label{eq:trace_form}
  \langle \Qhat_k \rangle = \Tr[\rhohat\,\Qhat_k]
  = \sum_{n=0}^\infty \rho_{nn}\, q_{kn}.
\end{equation}
This is an \emph{infinite dot product} of the occupation-probability vector
$\mathbf{p} = (\rho_{00}, \rho_{11}, \rho_{22}, \ldots)$ against the
eigenvalue vector $\mathbf{q}_k = (q_{k0}, q_{k1}, q_{k2},\ldots)$.

% ══════════════════════════════════════════════════════════════════════════════
\section{The Two Axioms}
\label{sec:axioms}
% ══════════════════════════════════════════════════════════════════════════════

\begin{axiom}[Flat Background]\label{ax:flat}
The pre-emergence arena is $(3{+}1)$-dimensional Minkowski spacetime:
$\eta_{\mu\nu} = \mathrm{diag}(-1,+1,+1,+1)$.
No curvature, no preferred time, no pre-existing matter.
\end{axiom}

\begin{remark}
Axiom~\ref{ax:flat} is foundational and self-standing.
It is not derived from general relativity, string theory, or any other
framework. Why four dimensions? Why Minkowski rather than another flat
geometry? These questions are not answered here; the axiom names the
background on which the theory is defined, and treats that choice as
primitive.
\end{remark}

\begin{axiom}[Global Zero-Sum Constraint]\label{ax:zs}
For all conserved charges $\Qhat_k$ and all times $t$:
\begin{equation}\label{eq:ZSC}
  \boxed{\Tr\bigl[\rhohat(t)\,\Qhat_k\bigr] = 0
  \quad \forall\, k,\; \forall\, t.}
\end{equation}
Equivalently, the observable charge vector
\begin{equation}\label{eq:charge_vector}
  \mathbf{Q}[\rhohat]
  = \begin{pmatrix}
    \Tr[\rhohat\,\Qhat_E] \\
    \Tr[\rhohat\,\Qhat_{P^{x}}] \\
    \Tr[\rhohat\,\Qhat_{P^{y}}] \\
    \Tr[\rhohat\,\Qhat_{P^{z}}] \\
    \Tr[\rhohat\,\Qhat_{L^{x}}] \\
    \Tr[\rhohat\,\Qhat_{L^{y}}] \\
    \Tr[\rhohat\,\Qhat_{L^{z}}] \\
    \Tr[\rhohat\,\Qhat_{\mathrm{em}}] \\
    \Tr[\rhohat\,\Qhat_{B-L}] \\
    \Tr[\rhohat\,\Qhat_{C_r}] \\
    \Tr[\rhohat\,\Qhat_{C_g}] \\
    \Tr[\rhohat\,\Qhat_{C_b}]
  \end{pmatrix}
  =
  \begin{pmatrix} 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{pmatrix}
  = \mathbf{0}_{12\times 1}.
\end{equation}
\end{axiom}

\begin{remark}
Energy (the Hamiltonian) is one conserved charge, associated with
time-translation symmetry by Noether's theorem.
Axiom~\ref{ax:zs} is therefore strictly stronger than requiring zero total
energy alone: it also sets net momentum, angular momentum, and every gauge
charge to zero.
The pre-emergence state is a quantum nothing in every sense that Noether's
theorem can assign a conserved number to.

Imposing zero energy alone would leave the possibility of a state with, for
example, nonzero angular momentum --- which would carry a preferred axis
into the emerged universe.
The full zero-sum axiom eliminates all such remnants simultaneously.
\end{remark}

\begin{remark}
Equation~\eqref{eq:ZSC} is an infinite linear constraint on $\rhohat$:
substituting \eqref{eq:trace_form}, it reads
$\sum_{n=0}^\infty \rho_{nn}\,q_{kn} = 0$ for each $k$.
This defines a hyperplane $\ZSC_k \subset \mathcal{B}(\Fock)$, and the
set of physically allowed density operators is
$\ZSC = \bigcap_k \ZSC_k \cap \{\text{density operators}\}$.
\end{remark}

\begin{remark}
Axiom~\ref{ax:flat} and Axiom~\ref{ax:zs} are the \emph{only} postulates
of TUO.
Every other result in this paper is derived from these two axioms plus
Standard Model particle content.
Where we invoke additional physical input, it is explicitly indicated.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Vacuum State and the Zero-Sum Filter}
\label{sec:vacuum}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{The vacuum}

\begin{definition}[Vacuum state]
The ground state prior to any excitation:
\begin{equation}
  |0\rangle = \begin{pmatrix}1\\0\\0\\\vdots\end{pmatrix},
  \qquad
  \rhohat_0 = |0\rangle\langle 0|
  = \begin{pmatrix}
    1 & 0 & 0 & \cdots \\
    0 & 0 & 0 & \cdots \\
    0 & 0 & 0 & \cdots \\
    \vdots & \vdots & \vdots & \ddots
  \end{pmatrix}.
\end{equation}
\end{definition}

\begin{lemma}
The vacuum state satisfies Axiom~\ref{ax:zs}.
\end{lemma}
\begin{proof}
$\Tr[\rhohat_0\Qhat_k] = q_{k0} = 0$ (vacuum carries no charge).
\end{proof}

\subsection{Single-particle prohibition}

\begin{theorem}\label{thm:single}
A pure single-particle state $|1;q\rangle$ with charge eigenvalue $q \neq 0$
violates Axiom~\ref{ax:zs}.
\end{theorem}
\begin{proof}
Let $|1;q\rangle = \begin{pmatrix}0\\1\\0\\\vdots\end{pmatrix}$,
$\rhohat_1 = |1;q\rangle\langle 1;q|$,
$\Qhat = \mathrm{diag}(0,q,0,\ldots)$.
Then $\Tr[\rhohat_1\Qhat] = q \neq 0$.
\end{proof}

\subsection{Pair configurations}

\begin{theorem}\label{thm:pair}
A superposition of states with opposite charge eigenvalues satisfies
Axiom~\ref{ax:zs} for that charge.
\end{theorem}
\begin{proof}
Let $|+\rangle = \begin{pmatrix}0\\1\\0\\\vdots\end{pmatrix}$,
$|-\rangle = \begin{pmatrix}0\\0\\1\\\vdots\end{pmatrix}$,
$|\psi\rangle = \frac{1}{\sqrt{2}}(|+\rangle + |-\rangle)$.
Then $\rhohat_\mathrm{pair} = |\psi\rangle\langle\psi| =
\frac{1}{2}\begin{pmatrix}0&0&\cdots\\0&1&1&\cdots\\0&1&1&\cdots\\
\vdots\end{pmatrix}$
and $\Qhat = \mathrm{diag}(0,+q,-q,\ldots)$, giving
$\Tr[\rhohat_\mathrm{pair}\Qhat] = \frac{1}{2}(q - q) = 0$.
\end{proof}

\begin{remark}
The proof works for any superposition of states whose charge eigenvalues
sum to zero.
The zero-sum constraint acts as a \emph{linear filter}: it admits any
state in the kernel of the charge functional, and rejects all others.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{The Complete Charge Equation}
\label{sec:full_balance}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Hamiltonian decomposition}

The total Hamiltonian splits as
$\Hhat = \Hhat_{\mathrm{matter}} + \Hhat_{\mathrm{field}}$,
where in the energy eigenbasis:
\begin{equation}
  \Hhat_{\mathrm{matter}} =
  \begin{pmatrix}
    0 & 0 & 0 & \cdots \\
    0 & E_1 & 0 & \cdots \\
    0 & 0 & E_2 & \cdots \\
    \vdots & \vdots & \vdots & \ddots
  \end{pmatrix},
  \quad E_n > 0.
\end{equation}

\begin{theorem}[Energy balance]\label{thm:energy_bal}
Axiom~\ref{ax:zs} applied to $\Hhat$ requires
\begin{equation}\label{eq:energy_balance}
  \Tr[\rhohat\,\Hhat_{\mathrm{matter}}]
  + \Tr[\rhohat\,\Hhat_{\mathrm{field}}] = 0,
\end{equation}
i.e., $E_{\mathrm{field}} = -E_{\mathrm{matter}}$ exactly.
\end{theorem}
\begin{proof}
$\Tr[\rhohat\,\Hhat] = 0 = \Tr[\rhohat(\Hhat_\mathrm{m}+\Hhat_\mathrm{f})]
 = E_\mathrm{m} + E_\mathrm{f}$.
\end{proof}

\subsection{The complete zero equation}

All 12 conserved-charge observables form a column vector.
Axiom~\ref{ax:zs} states that this vector equals the zero column:
\begin{equation}\label{eq:full_zero}
  \underbrace{
  \begin{pmatrix}
    \Tr[\rhohat\,\Qhat_E] \\[2pt]
    \Tr[\rhohat\,\Qhat_{P^{\mu}}] \\[2pt]
    \Tr[\rhohat\,\Qhat_{J}] \\[2pt]
    \Tr[\rhohat\,\Qhat_{\mathrm{em}}] \\[2pt]
    \Tr[\rhohat\,\Qhat_{B-L}] \\[2pt]
    \Tr[\rhohat\,\Qhat_{\mathbf{C}}]
  \end{pmatrix}}_{\mathbf{Q}[\rhohat]\;\in\;\mathbb{R}^{12}}
  \;=\;
  \begin{pmatrix}0 \\ \mathbf{0}_4 \\ \mathbf{0}_3 \\ 0 \\ 0 \\ \mathbf{0}_3 \end{pmatrix}
  \;=\;\mathbf{0}_{12\times 1}.
\end{equation}
This is the \emph{central equation of TUO}:
the universe's full observable content equals the infinite zero.

\begin{remark}
The zero on the right-hand side is not an approximation.
It is exact and holds at all times by Axiom~\ref{ax:zs}.
The field contribution is whatever it must be to maintain the balance:
for each matter configuration with charges $\mathbf{Q}_\mathrm{m}$, the
field sector must carry $\mathbf{Q}_\mathrm{f} = -\mathbf{Q}_\mathrm{m}$.
\end{remark}

\subsection{Time evolution preserves zero-sum}

\begin{theorem}\label{thm:liouville}
Unitary evolution $\rhohat(t) = \hat{U}(t)\rhohat(0)\hat{U}^\dagger(t)$
preserves Axiom~\ref{ax:zs} if and only if $[\Hhat, \Qhat_k] = 0$ for all $k$.
\end{theorem}
\begin{proof}
$\frac{d}{dt}\Tr[\rhohat\Qhat_k]
 = \Tr\!\left[\frac{\partial\rhohat}{\partial t}\Qhat_k\right]
 = -\frac{i}{\hbar}\Tr[[\Hhat,\rhohat]\Qhat_k]
 = -\frac{i}{\hbar}\Tr[\rhohat[\Qhat_k,\Hhat]]$.
This vanishes if and only if $[\Hhat,\Qhat_k] = 0$, i.e., each $\Qhat_k$
is conserved.
\end{proof}

% ══════════════════════════════════════════════════════════════════════════════
\section{The Selection Problem: Why Generic Fluctuations Cannot Become a Universe}
\label{sec:selection}
% ══════════════════════════════════════════════════════════════════════════════

Axiom~\ref{ax:zs} permits a vast class of zero-sum fluctuations from the
vacuum. Theorems~\ref{thm:single} and~\ref{thm:pair} show that the constraint
acts as a filter: single charged particles are excluded, but any superposition
whose charges cancel is admitted. The vacuum can fluctuate. The question is
which fluctuations can survive long enough, and carry enough energy, to become
a universe. Two explicit barriers prevent generic fluctuations from doing so.

\subsection{Barrier I: The energy gap}

A typical virtual pair --- an electron and positron --- carries energy
\begin{equation}
  E_{e^+e^-} = 2m_e c^2 \approx 1.64 \times 10^{-13}\;\mathrm{J}.
\end{equation}
The energy required to seed a Planck-scale cell capable of expanding into
a universe is (Section~\ref{sec:energy}):
\begin{equation}
  E_{\mathrm{cell}} = \gstar \cdot \frac{\EPl}{2} \approx 1.04 \times 10^{11}\;\mathrm{J}.
\end{equation}
The ratio is
\begin{equation}
  \frac{E_{e^+e^-}}{E_{\mathrm{cell}}} \approx 1.6 \times 10^{-24}.
\end{equation}
A generic virtual pair carries twenty-four orders of magnitude too little
energy. No accumulation mechanism exists within a single Planck-time window
to bridge this gap from below.

\subsection{Barrier II: The annihilation timescale}

The Heisenberg time-energy relation constrains the lifetime of any virtual
configuration:
\begin{equation}
  \Delta t \sim \frac{\hbar}{\Delta E}.
\end{equation}
At Planck energies, $\Delta t \sim \tPl \approx 5.4 \times 10^{-44}$~s.
Within this window, a generic fluctuation appears and annihilates before
any complex structure can form. Even if an arbitrary collection of
particle-antiparticle pairs were to emerge, each pair presents its own
annihilation channel --- $e^+e^- \to \gamma\gamma$, $q\bar{q} \to gg$,
and so on --- and the configuration collapses back to the vacuum on the
Planck timescale.

\subsection{The resolution: the maximum fluctuation}

The two barriers are simultaneously and uniquely overcome by the most extreme
fluctuation consistent with Axiom~\ref{ax:zs}: \emph{all Standard Model
degrees of freedom emerging at the same point in space at the same moment
in time}. Not sequentially. Not spread out. All of them, simultaneously,
co-located at a single Planck-scale volume.

Concretely, this is a configuration in which all $\gstar = 106.75$
effective SM degrees of freedom are excited to their Heisenberg minimum
energy $\EPl/2$, co-located in one Planck cell, simultaneously.

\begin{theorem}[Maximum fluctuation uniquely overcomes both barriers]
\label{thm:maxfluc}
The configuration consisting of all SM matter --- quarks and leptons across
three generations, in all colours and flavours --- simultaneously co-located
at one Planck cell, with no antiparticles:
\begin{enumerate}[label=(\roman*)]
  \item satisfies Axiom~\ref{ax:zs}, and
  \item carries energy $E_{\mathrm{cell}} = \gstar \cdot \EPl/2$, and
  \item has no annihilation channel and is therefore stable.
\end{enumerate}
\end{theorem}
\begin{proof}
\textit{(i) Zero-sum satisfied.}
The SM fermion content satisfies $B - L = 0$ and $Q = 0$ per generation
(Theorem~\ref{thm:bmL} below); the total vanishes for all three generations.
Gravitational energy $E_\mathrm{grav} = -E_\mathrm{cell}$ by
Theorem~\ref{thm:energy_bal}. All twelve charges in
Eq.~\eqref{eq:charge_vector} vanish.

\textit{(ii) Energy.}
Each of the $\gstar$ modes carries $\EPl/2$ from the Heisenberg bound.
Total: $E_{\mathrm{cell}} = \gstar \cdot \EPl/2$ (Theorem~\ref{thm:energy}
below).

\textit{(iii) Stability.}
The configuration contains only matter: antiparticle occupation numbers
$n_{\bar{f}} = 0$ for all $\bar{f}$.
By Theorem~\ref{thm:stability} below, no SM process can open an annihilation
channel. The configuration is kinematically stable.
\end{proof}

\begin{remark}
The resolution is not a timescale argument. It is a \emph{kinematic} one.
Annihilation requires a particle and its antiparticle. The maximum
fluctuation contains no antiparticles. There is no annihilation, not
because the timescale is too short, but because the required reactant is
absent by construction. This is what the No-Annihilation Theorem
(Theorem~\ref{thm:stability}) will make precise.
\end{remark}

\begin{remark}
This is the genuine creative core of TUO.
The zero-sum vacuum is not inert: it fluctuates.
Generically its fluctuations are too small and too short-lived --- the two
barriers above are absolute.
The unique exception is the configuration that carries all SM content,
simultaneously, co-located: it is energetically sufficient and kinematically
stable.
It is the only zero-sum fluctuation the vacuum can produce that is capable
of becoming a universe.
The number $\gstar = 106.75$ enters not as a free parameter, but as the
total count of SM degrees of freedom that must all emerge together for
either barrier to be overcome.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Configuration Space and Frozen Constants}
\label{sec:configs}
% ══════════════════════════════════════════════════════════════════════════════

The maximum fluctuation selects \emph{a} configuration, not necessarily
\emph{the only} configuration. This section characterises the full space
of zero-sum configurations that can produce stable universes, and identifies
which aspects of physical law are fixed by the axioms versus fixed by the
particular configuration that emerged.

\subsection{The space of stable configurations}

A configuration is a choice of field content (determining $\gstar$),
charge assignments (determining whether Axiom~\ref{ax:zs} is satisfiable
with no antiparticles), and a mass spectrum.
The zero-sum filter admits a matter-only configuration if and only if
it satisfies the anomaly-cancellation conditions: $Q = 0$ and $B - L = 0$
per generation-like block.
These are identical to the $\mathrm{SU}(5)$ grand-unification consistency
conditions, which the Standard Model satisfies exactly
(Theorem~\ref{thm:bmL}).

Other configurations with $N_\mathrm{gen} \neq 3$ also satisfy these
conditions, since the anomaly-cancellation identity holds
\emph{per generation} independently.
The corresponding $\gstar$ values and cell energies are shown in
Table~\ref{tab:configs}.

\begin{table}[ht]
\centering
\caption{Stable zero-sum configurations for varying number of generations.
Each configuration satisfies $Q = 0$, $B - L = 0$, passes the no-annihilation
test, and could in principle become a universe with the given cell energy.
$N_\nu$ is measured at LEP~\cite{PDG2022}: only $N_\mathrm{gen} = 3$ is
consistent with observation.}
\label{tab:configs}
\begin{tabular}{cccc}
\toprule
$N_\mathrm{gen}$ & $\gstar$ & $E_\mathrm{cell}/\EPl$ & $E_\mathrm{cell}$ (J) \\
\midrule
1 & 42.00 & 21.00 & $4.11 \times 10^{10}$ \\
2 & 56.00 & 28.00 & $5.48 \times 10^{10}$ \\
\textbf{3} & \textbf{70.00} & \textbf{35.00} & $\mathbf{6.85 \times 10^{10}}$ \\
4 & 84.00 & 42.00 & $8.22 \times 10^{10}$ \\
5 & 98.00 & 49.00 & $9.58 \times 10^{10}$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}
Note: the $\gstar$ values in Table~\ref{tab:configs} count only fermion
generations; the bosonic sector ($N_B = 28$ for the SM gauge and Higgs
bosons) may also differ across configurations.
The value $\gstar = 106.75$ in Table~\ref{tab:predictions} includes the
full SM boson content.
\end{remark}

\begin{remark}
Each row of Table~\ref{tab:configs} describes a logically consistent
configuration that satisfies both axioms. They correspond to different
possible universes. The observation $N_\nu = 2.9840 \pm 0.0082$ (LEP)
selects $N_\mathrm{gen} = 3$ from among these possibilities.
TUO does not derive $N_\mathrm{gen} = 3$; this is an open problem
(Section~\ref{sec:open}).
\end{remark}

\subsection{What freezes at emergence: a classification}

When a configuration emerges, three classes of quantities are fixed.

\textbf{Class I --- Axiom-fixed constants.}
The speed of light $c$ is fixed by Axiom~\ref{ax:flat}: Minkowski spacetime
has a unique light-cone structure, and $c$ is its invariant speed.
The reduced Planck constant $\hbar$ is required for the Heisenberg bound
in Axiom~\ref{ax:zs}. Newton's constant $G$ is required by
Theorem~\ref{thm:energy_bal}: the gravitational energy $E_\mathrm{grav} = -G\MPl^2/(2\lPl)$
must exactly cancel $E_\mathrm{matter} = \EPl/2$, which forces
$G\MPl^2 = \hbar c$ --- an identity that defines $G$ in terms of
$\hbar$ and $c$. These three constants are \emph{not} configuration-dependent;
they are consequences of the axioms themselves.

\textbf{Class II --- Configuration-fixed dimensionless ratios.}
The fine-structure constant $\alpha \approx 1/137$, the strong coupling
$\alpha_s$, and the mass ratios $m_f/\MPl$ are fixed when a particular
configuration is selected. A universe with $N_\mathrm{gen} = 4$ would,
in general, have different values of these ratios. These are the parameters
of the effective Lagrangian and they vary across configurations.

\textbf{Class III --- Emergent effective constants.}
The cosmological constant $\Lambda$, the Higgs vacuum expectation value,
and the QCD scale $\Lambda_\mathrm{QCD}$ are dynamically determined within
a configuration. They are consequences of Class~II, not independent choices.

\begin{remark}
A common intuition is that the speed of light could be different in another
universe. TUO refines this: $c$ as a \emph{speed limit} cannot vary because
it is fixed by Axiom~\ref{ax:flat}. What can vary is the ratio $m_e c^2 / \EPl$
--- the electron is lighter or heavier relative to the Planck scale ---
which would alter all of chemistry, not the geometry of spacetime.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Standard Model Content and the $B-L$ Constraint}
\label{sec:sm}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Why $B - L$, not $B$ alone}

TUO constrains $B - L = 0$ rather than $B = 0$ separately.
The constraint $B = 0$ would require equal numbers of quarks and antiquarks,
forbidding matter-only universes.
The constraint $B - L = 0$ permits matter-only configurations in which
$B = L \neq 0$, provided their difference vanishes.

\begin{theorem}\label{thm:bmL}
The Standard Model three-generation fermion content satisfies
$Q = 0$ and $B - L = 0$ per generation as an algebraic identity.
\end{theorem}
\begin{proof}
In generation $g$, denote the quark doublet $(u_g, d_g)$ in three colour
copies and the lepton doublet $(\nu_g, e_g)$.
\medskip\\
\textit{Baryon minus lepton number:}\\
$B_g = 2\,\text{flavours} \times 3\,\text{colours} \times \tfrac{1}{3} = 2$;
$L_g = 2\,\text{leptons} \times 1 = 2$;
therefore $B_g - L_g = 0$.
\medskip\\
\textit{Electric charge:}
\[
  Q_g
  = 3\!\left(\tfrac{2}{3}\right) + 3\!\left(-\tfrac{1}{3}\right) + (-1) + 0
  = 2 - 1 - 1 = 0.
\]
Both identities hold for $g = 1, 2, 3$ by the same counting.
\end{proof}

\begin{remark}
This result is the anomaly-cancellation condition of $\mathrm{SU}(5)$
grand unification.
TUO does not derive the SM content; rather, the SM's structure satisfies
TUO's constraint.
Whether TUO can derive the SM content from first principles remains an
open question (Section~\ref{sec:open}).
\end{remark}

\subsection{Matter-only stability}

\begin{theorem}[No-Annihilation]\label{thm:stability}
A configuration satisfying $Q = 0$, $B-L = 0$, and containing no
antiparticles has no SM annihilation channel.
\end{theorem}
\begin{proof}
Every SM annihilation reaction requires a particle-antiparticle pair:
$e^- e^+ \to \gamma\gamma$, $q\bar{q} \to gg$, etc.
In a configuration with zero antiparticle occupation ($n_{\bar{f}} = 0$ for
all $\bar{f}$), no such pair exists.
No SM process can produce $\bar{f}$ from an initial state with $n_{\bar{f}}=0$
without violating lepton or baryon number conservation.
\end{proof}

\begin{remark}
The argument is purely kinematic: the absence of antiparticles removes
the kinematic possibility of annihilation.
No claim is made about interaction timescales or thermodynamic equilibrium.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Energy from Heisenberg Uncertainty}
\label{sec:energy}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{The energy matrix at $t = \tPl$}

The Heisenberg time-energy uncertainty relation gives the minimum energy
per mode at timescale $\Delta t = \tPl$:
\begin{equation}\label{eq:heisenberg}
  \Delta E\,\Delta t \geq \frac{\hbar}{2}
  \;\implies\;
  \Delta E_{\min} = \frac{\hbar}{2\tPl} = \frac{\EPl}{2}.
\end{equation}
At $T = \TPl$, all SM fields are relativistic.
The number of relativistic degrees of freedom is
\begin{equation}\label{eq:gstar}
  \gstar = N_B + \tfrac{7}{8}N_F
  = 28 + \tfrac{7}{8}(90) = 106.75,
\end{equation}
where $N_B = 28$ (gauge and Higgs bosons) and $N_F = 90$ (fermionic
polarisations and colour states).

The Hamiltonian for the emergence configuration, restricted to the
first $\gstar$ excitation modes at their minimum Heisenberg energies,
takes the form:
\begin{equation}\label{eq:H_emergence}
  \Hhat_{\mathrm{emergence}}
  = \begin{pmatrix}
    0 & & & & \\
    & \frac{\EPl}{2} & & & \\
    & & \frac{\EPl}{2} & & \\
    & & & \ddots & \\
    & & & & \frac{\EPl}{2}
  \end{pmatrix}_{\gstar \times \gstar}
  \oplus \mathbf{0}_{\infty-\gstar},
\end{equation}
with $\gstar$ copies of $\EPl/2$ on the diagonal.

\begin{theorem}[Energy per Planck cell]\label{thm:energy}
The total matter energy per Planck cell is
\begin{equation}\label{eq:Ecell_total}
  \boxed{E_{\mathrm{cell}} = \gstar \cdot \frac{\EPl}{2} = 53.375\,\EPl.}
\end{equation}
By Theorem~\ref{thm:energy_bal}, the gravitational field energy satisfies
$E_{\mathrm{grav}} = -E_{\mathrm{cell}}$ exactly.
The total energy of the universe is zero.
\end{theorem}
\begin{proof}
Each of the $\gstar$ modes carries minimum energy $\EPl/2$ from
Eq.~\eqref{eq:heisenberg}.
The sum is immediate.
$E_{\mathrm{grav}} = -E_{\mathrm{cell}}$ follows from Theorem~\ref{thm:energy_bal}
applied to $\Hhat_\mathrm{field} = \Hhat_\mathrm{grav}$.
\end{proof}

\subsection{Comparison with standard cosmology}

The energy density of a thermal plasma at $T = \TPl$ in volume $\lPl^3$,
from the Standard Model equation of state, is
\begin{equation}\label{eq:E_std}
  E_{\mathrm{std}}
  = \frac{\pi^2}{30}\,\gstar\,\frac{(k_B\TPl)^4}{(\hbar c)^3}\,\lPl^3
  = 6.87 \times 10^{10}\;\mathrm{J}.
\end{equation}
The TUO value is $E_\mathrm{cell} = 53.375\,\EPl = 1.044 \times 10^{11}\;\mathrm{J}$.
Their ratio is
\begin{equation}\label{eq:ratio}
  \frac{E_{\mathrm{cell}}}{E_{\mathrm{std}}} = \frac{15}{\pi^2} \approx 1.52.
\end{equation}
This factor is a pure dimensionless number independent of all SM parameters.
It is the ratio of the Heisenberg minimum energy to the thermal average energy
per mode.

% ══════════════════════════════════════════════════════════════════════════════
\section{Simultaneous Emergence: A Possible Mechanism}
\label{sec:path_integral}
% ══════════════════════════════════════════════════════════════════════════════

\begin{remark}
The analysis in this section is a \emph{possible mechanism}, not a proven
theorem. The saddle-point argument below suggests that simultaneous
emergence of all Planck cells is exponentially preferred over localised
fluctuations under the constrained measure, but a rigorous proof --- with
controlled error bounds on the saddle-point approximation and a
connection to the observed $\Delta T/T \sim 10^{-5}$ --- is left as
Open Question~\ref{oq:cmb}.
\end{remark}

\subsection{Setup}

Partition the Minkowski background (Axiom~\ref{ax:flat}) into a lattice of
Planck cells indexed by $i = 1,\ldots,N$.
The number of cells in the observable universe's comoving volume at
$t = \tPl$ is
\begin{equation}\label{eq:Ncells}
  N = \left(\frac{R_0\sqrt{\tPl/t_0}}{\lPl}\right)^3 \approx 8.8 \times 10^{92}.
\end{equation}
The constrained Euclidean path integral is
\begin{equation}\label{eq:PI}
  Z = \int \mathcal{D}\phi\;e^{-S_E[\phi]}
  \prod_k \delta\!\bigl(Q_k[\phi]\bigr),
\end{equation}
where the delta functions enforce Axiom~\ref{ax:zs} globally.

\subsection{Saddle-point analysis}

Writing $\delta(Q_k) = \int \frac{d\lambda_k}{2\pi}e^{i\lambda_k Q_k}$:
\begin{equation}\label{eq:PI_lambda}
  Z = \int \mathcal{D}\phi\,\mathcal{D}\lambda\;
  \exp\!\left(-S_E[\phi] + i\sum_k\lambda_k Q_k[\phi]\right).
\end{equation}

At the saddle point, single-cell fluctuations against $N-1$ vacuum cells
incur an action penalty $\Delta S \sim N$, while simultaneous $N$-cell
emergence satisfies the global constraint with $\Delta S = 0$.
The ratio of partition-function contributions is formally
\begin{equation}
  \frac{Z_{\mathrm{simultaneous}}}{Z_{\mathrm{single}}}
  \sim e^{N} = e^{8.8 \times 10^{92}}.
\end{equation}
This suggests simultaneous emergence is the overwhelmingly dominant
configuration, offering a possible explanation for the observed
uniformity of the CMB.
A full proof, controlling the saddle-point approximation and connecting
the amplitude to the observed $\Delta T/T$, is left as
Open Question~\ref{oq:cmb}.

% ══════════════════════════════════════════════════════════════════════════════
\section{Expansion Law from Quantum Wavepacket Spreading}
\label{sec:expansion}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Wavepacket initial condition}

At $t = \tPl$, the emergent configuration is localised to one Planck cell:
\begin{equation}\label{eq:psi0}
  \psi(\mathbf{x},0)
  = (2\pi\lPl^2)^{-3/4}\exp\!\left(-\frac{x^2}{4\lPl^2}\right).
\end{equation}
Position-momentum uncertainty with $\Delta x = \lPl$:
\begin{equation}
  \Delta p \geq \frac{\hbar}{2\lPl} = \frac{\MPl c}{2}.
\end{equation}
Relativistic velocity from $E = \EPl/2$, $p = \MPl c/2$:
\begin{equation}\label{eq:v_c}
  v_0 = \frac{pc^2}{E} = \frac{(\MPl c/2)c^2}{\EPl/2} = c.
\end{equation}
This immediately gives $w = P_\mathrm{kinetic}/(uc^2) = 1/3$ (radiation),
which TUO derives rather than assumes.

\subsection{Derived expansion law}

\begin{theorem}[Expansion law]\label{thm:expansion}
For massless particles ($v = c$), the wavepacket width evolves as
\begin{equation}\label{eq:sigma}
  \sigma(t) = \lPl\sqrt{1 + \left(\frac{ct}{\lPl}\right)^2},
\end{equation}
giving the volume
\begin{equation}\label{eq:V_t}
  V(t) = \frac{4\pi}{3}\,\sigma(t)^3
  = \frac{4\pi}{3}\,\lPl^3\!\left[1 + \left(\frac{ct}{\lPl}\right)^2\right]^{3/2}.
\end{equation}
For $ct \gg \lPl$:
\begin{equation}\label{eq:V_classical}
  V(t) \approx \frac{4\pi}{3}(ct)^3
  \left[1 + \frac{\Delta V}{V}\right],
  \quad
  \frac{\Delta V}{V} = 3\!\left(\frac{\lPl}{ct}\right)^2.
\end{equation}
\end{theorem}

\begin{proposition}[Sub-luminal bound]\label{prop:sub}
$v(t) = d\sigma/dt < c$ for all $t > 0$.
\end{proposition}
\begin{proof}
$v(t) = c(ct/\lPl)/\sqrt{1+(ct/\lPl)^2} < c$.
\end{proof}

\begin{remark}
The expansion speed in TUO is sub-luminal at all times.
Standard inflation requires super-luminal expansion of the scale factor.
This is a distinguishing feature of TUO.
\end{remark}

\subsection{Hubble parameter}

\begin{equation}\label{eq:H_TUO}
  H_{\mathrm{TUO}}(t)
  = \frac{1}{\sigma}\frac{d\sigma}{dt}
  = \frac{c^2 t}{\lPl^2 + c^2 t^2}.
\end{equation}
At $t = \tPl$, using $\lPl = c\tPl$:
\begin{equation}\label{eq:H_junction}
  H_{\mathrm{TUO}}(\tPl)
  = \frac{c^2\tPl}{2c^2\tPl^2}
  = \frac{1}{2\tPl}
  = H_{\mathrm{FRW}}(\tPl).
\end{equation}
The Hubble parameter is continuous across the TUO--FRW junction.

% ══════════════════════════════════════════════════════════════════════════════
\section{Junction Conditions at $t = \tPl$}
\label{sec:junction}
% ══════════════════════════════════════════════════════════════════════════════

\begin{theorem}[Smooth junction]\label{thm:junction}
The TUO solution at $t = \tPl^-$ matches the FRW radiation-dominated
solution at $t = \tPl^+$ with no discontinuity in $H$, $w$, or $k$.
\end{theorem}
\begin{proof}
Three conditions must be checked:
\begin{enumerate}[label=(\roman*)]
  \item \textit{Hubble parameter}: verified in Eq.~\eqref{eq:H_junction}. \checkmark
  \item \textit{Equation of state}: TUO gives $w = 1/3$ from
        Eq.~\eqref{eq:v_c}; FRW takes $w = 1/3$ as input.
        TUO derives what FRW assumes. \checkmark
  \item \textit{Spatial curvature}: Axiom~\ref{ax:flat} requires $k = 0$
        exactly; FRW allows $k = 0$; Planck 2018 measures
        $\Omega = 1.0007 \pm 0.0037$~\cite{Planck2018}. \checkmark
\end{enumerate}
No surface energy term is required.
\end{proof}

\begin{remark}
The energy density comparison (junction condition~IV) gives a mismatch
by a factor $15/\pi^2 \approx 1.52$.
This does not break the smooth junction for $H$, $w$, $k$; it reflects
the difference between the TUO Heisenberg minimum-energy state and the
FRW thermal-equilibrium state.
The physical interpretation of this factor is left as an open problem.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Thermal History}
\label{sec:thermal}
% ══════════════════════════════════════════════════════════════════════════════

\begin{theorem}[Pre-emergence temperature]\label{thm:T_TUO}
The temperature $T_{\mathrm{TUO}}$ of a radiation bath with $\gstar$ degrees of
freedom having the same energy density as the TUO pre-emergence state is
\begin{equation}
  T_{\mathrm{TUO}} = \left(\frac{15}{\pi^2}\right)^{1/4}\TPl \approx 1.11\,\TPl,
\end{equation}
independent of $\gstar$.
\end{theorem}
\begin{proof}
Set $\rho_\mathrm{TUO} = \gstar \cdot \EPl/(2\lPl^3)$ equal to the
Stefan-Boltzmann density $\rho_\mathrm{SB} = (\pi^2/30)\gstar(k_B T)^4/(\hbar c)^3$.
The factor $\gstar$ cancels on both sides, yielding
$(k_B T)^4 = (15/\pi^2)(k_B\TPl)^4$.
\end{proof}

\begin{remark}
The $\gstar$-independence is non-trivial: the particle content determines the
total energy but not the temperature, because both the Heisenberg energy and
the Stefan-Boltzmann density scale identically with $\gstar$.
The temperature after the junction follows the standard radiation-dominated law
$T(t) = T_{\mathrm{TUO}}\sqrt{\tPl/t}$.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Equality in Fundamental Equations as a Consequence of the Zero-Sum Axiom}
\label{sec:equalities}
% ══════════════════════════════════════════════════════════════════════════════

A striking feature of fundamental physics is that its governing equations
are \emph{equalities}: $G_{\mu\nu} = \kappa T_{\mu\nu}$, not $G_{\mu\nu}
\geq \kappa T_{\mu\nu}$. This is not obvious. Why should geometry exactly
balance matter-energy, rather than exceed or fall short of it? TUO provides
a structural answer.

\subsection{Why equalities, not inequalities}

Axiom~\ref{ax:zs} states $\Tr[\rhohat\,\Qhat_k] = 0$ \emph{exactly} for
every conserved charge.
This propagates: for any local density $\mathcal{Q}_k(x)$ associated with
$\Qhat_k$,
\begin{equation}
  \int \mathcal{Q}_k(x)\,d^3x = 0 \quad \forall\, t.
\end{equation}
If at any spacetime point $x_0$ the local charge density were positive,
Axiom~\ref{ax:zs} requires it to be exactly compensated elsewhere.
Localising this balance at every point requires
\begin{equation}
  \mathcal{Q}_k^\mathrm{matter}(x) + \mathcal{Q}_k^\mathrm{field}(x) = 0
  \quad \forall\, x.
\end{equation}
This is an equality because any surplus or deficit at a single point
violates the axiom. The sign $=$ is not empirical; it is forced.

\subsection{The Einstein field equations}

Applied to the energy-momentum charge, Theorem~\ref{thm:energy_bal} gives
$E_\mathrm{matter} + E_\mathrm{grav} = 0$ globally. Localising yields
$T_{\mu\nu}^\mathrm{matter}(x) + T_{\mu\nu}^\mathrm{grav}(x) = 0$ pointwise.
By the Lovelock theorem~\cite{Lovelock1971}, the unique rank-2 symmetric
divergence-free tensor constructible from the metric alone is
$aG_{\mu\nu} + bg_{\mu\nu}$; matching the Newtonian limit fixes
$a = 8\pi G/c^4$. One therefore has:
\begin{equation}\label{eq:EFE_ZSC}
  G_{\mu\nu} + \Lambda g_{\mu\nu} = \frac{8\pi G}{c^4}\,T_{\mu\nu}.
\end{equation}
The left side is the gravitational charge density; the right side is the
matter charge density. Their sum is zero: this is the Einstein field
equation, and its equality sign is the local expression of Axiom~\ref{ax:zs}.
TUO does not \emph{derive} the Lovelock theorem; rather, given that theorem,
the EFE structure is \emph{forced} by the axiom.

\subsection{Maxwell's equations}

For the electromagnetic charge $Q_\mathrm{em}$, the local zero-sum reads
$\partial_\mu F^{\mu\nu}(x) + J^\nu(x)/\varepsilon_0 = 0$,
i.e.,
\begin{equation}
  \partial_\mu F^{\mu\nu} = \frac{J^\nu}{\varepsilon_0}.
\end{equation}
The field exactly compensates its source. An inequality would mean the
electromagnetic field carries more (or less) charge than the source creates,
violating Axiom~\ref{ax:zs}.

\subsection{The Schr\"odinger and Dirac equations}

For probability conservation ($Q_\mathrm{prob} = \int |\psi|^2 d^3x = 1$),
Axiom~\ref{ax:zs} requires this integral to be constant: $d/dt \int |\psi|^2
d^3x = 0$. The unique evolution that achieves this while being linear and
first-order in time is $i\hbar\,\partial_t\psi = \Hhat\psi$. Similarly, the
Dirac equation $(i\gamma^\mu\partial_\mu - m)\psi = 0$ enforces zero-sum
between positive and negative spinor components.

\begin{remark}
The structure of these observations is the same in each case: Axiom~\ref{ax:zs}
says a conserved charge totals to zero; the corresponding dynamical equation
is the local expression of that balance. The equality sign in every fundamental
equation is the local expression of $\mathbf{Q}[\rhohat] = \ZeroMat$.
\end{remark}

\subsection{Exotic phenomena permitted by the zero-sum axiom}

The axiom constrains but does not over-constrain. Table~\ref{tab:exotic}
lists phenomena that satisfy Axiom~\ref{ax:zs} and are therefore admitted
--- some observed, some conjectural.

\begin{table}[ht]
\centering
\caption{Phenomena permitted by the zero-sum axiom. Status:
$\checkmark$ = observed; $\sim$ = conjectural direction consistent with ZSC.}
\label{tab:exotic}
\begin{tabular}{p{3.2cm}p{6.8cm}c}
\toprule
Phenomenon & Zero-sum mechanism & Status \\
\midrule
Virtual particles &
  Particle-antiparticle pair: $Q_\mathrm{pair}=0$. Heisenberg limits
  duration, not existence. & $\checkmark$ \\
Quantum entanglement &
  Singlet state enforces $\langle S_z\rangle_A + \langle S_z\rangle_B = 0$
  at all separations (Section~\ref{sec:thermal}). & $\checkmark$ \\
CPT symmetry &
  Every process that creates particles has a CPT-conjugate creating
  antiparticles; global $B-L=0$ is maintained. & $\checkmark$ \\
Black hole information &
  Information (von Neumann entropy) is a conserved charge; ZSC requires
  $S_\mathrm{matter} + S_\mathrm{grav} = \mathrm{const}$: information
  transfers to the gravitational sector, not destroyed. & $\sim$ \\
Dark sector &
  Any hidden sector with internal $Q_\mathrm{dark}=0$,
  $B_\mathrm{dark}-L_\mathrm{dark}=0$ satisfies ZSC.
  Dark matter as a parallel zero-sum sector is admitted. & $\sim$ \\
Parallel configurations &
  Other $N_\mathrm{gen}$ configurations (Table~\ref{tab:configs}) satisfy ZSC
  and could in principle produce stable universes with
  different dimensionless ratios. & $\sim$ \\
\bottomrule
\end{tabular}
\end{table}


% ══════════════════════════════════════════════════════════════════════════════
\section{Statistical Analysis of the Energy Result}
\label{sec:stats}

The energy per Planck cell derived in Theorem~\ref{thm:energy},
\begin{equation}
  E_\mathrm{cell} = \frac{\gstar}{2}\,\EPl = 53.375\,\EPl,
\end{equation}
follows from two axioms and the Heisenberg bound, with no free parameters
beyond the SM particle count $\gstar = 106.75$.
This section analyses how extraordinary this result is.

\subsection{The universal factor $15/\pi^2$}

Setting the TUO energy density equal to the Stefan-Boltzmann density at
the same temperature, the factor $\gstar$ cancels (Theorem~\ref{thm:T_TUO}),
leaving:
\begin{equation}
  \frac{E_\mathrm{cell}}{E_\mathrm{thermal}}
  = \frac{\gstar \cdot \EPl/2}
         {(\pi^2/30)\,\gstar\,(k_B\TPl)^4/(\hbar c)^3 \cdot \lPl^3}
  = \frac{15}{\pi^2}
  \approx 1.5198.
\end{equation}
The ratio $15/\pi^2$ is a pure dimensionless number independent of:
\begin{enumerate}[leftmargin=2.5em,label=\textbullet]
  \item All SM coupling constants ($\alpha$, $\alpha_s$, $\alpha_w$)
  \item All particle masses ($m_e$, $m_u$, $m_H$, \ldots)
  \item All cosmological parameters ($H_0$, $\Omega$, $\Lambda$, \ldots)
  \item The number of degrees of freedom $\gstar$ (it cancels exactly)
\end{enumerate}
Its sole origin is the ratio of the Heisenberg minimum energy
($E_\mathrm{min} = \hbar/2t_\mathrm{Pl}$, giving the numerator $15$)
to the Stefan-Boltzmann radiation constant ($\pi^2/30$, giving the
denominator $\pi^2$).
This is a \emph{structural} result: it holds for any theory whose
pre-emergence state saturates the Heisenberg bound.

\subsection{Comparison of derivation inputs}

Standard cosmology arrives at the Planck-era energy density via the
FRW equation, SM thermodynamics, the measured values of $H_0$, $\Omega_b$,
$\Omega_r$, and approximately 20 additional parameters.
TUO arrives at $E_\mathrm{cell}$ from:
\begin{center}
\begin{tabular}{ll}
\toprule
Input & Status \\
\midrule
Axiom I (Minkowski background) & Postulate \\
Axiom II (zero-sum, $\Tr[\rhohat\Qhat_k]=0$) & Postulate \\
Heisenberg uncertainty $\Delta E\,\Delta t \geq \hbar/2$ & Standard QM \\
SM effective DOF count $\gstar = 106.75$ & From experiment \\
\bottomrule
\end{tabular}
\end{center}
No dynamical equations, no measured cosmological parameters, and no
initial conditions beyond the axioms are required.

\subsection{Energy scale in context}

Table~\ref{tab:energy_context} places $E_\mathrm{cell}$ in the hierarchy
of known energy scales. The result derived from two axioms sits
at the intersection of quantum mechanics (the Heisenberg bound) and
the Planck scale, and exceeds every laboratory energy scale by many orders
of magnitude.

\begin{table}[ht]
\centering
\caption{Energy scale comparisons. $E_\mathrm{cell} = 1.044 \times 10^{11}$~J is
derived from two axioms alone, with $\gstar = 106.75$ as the only experimental input.}
\label{tab:energy_context}
\begin{tabular}{lcc}
\toprule
Reference scale & Energy (J) & $E/E_\mathrm{cell}$ \\
\midrule
Electron rest mass $m_e c^2$ & $8.19 \times 10^{-14}$ & $7.8 \times 10^{-25}$ \\
Proton rest mass $m_p c^2$ & $1.50 \times 10^{-10}$ & $1.4 \times 10^{-21}$ \\
LHC full beam energy & $3.15 \times 10^{-3}$ & $3.0 \times 10^{-14}$ \\
1 megaton nuclear yield & $4.0 \times 10^{15}$ & $3.8 \times 10^{4}$ \\
\textbf{$E_\mathrm{cell}$ (TUO, from axioms)} & $\mathbf{1.044 \times 10^{11}}$ & $\mathbf{1}$ \\
Planck energy $\EPl$ & $1.956 \times 10^{9}$ & $1/53.4$ \\
Solar mass-energy $M_\odot c^2$ & $1.79 \times 10^{47}$ & $1.7 \times 10^{36}$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Significance of the zero-total-energy result}

The result $E_\mathrm{matter} + E_\mathrm{grav} = 0$ (Theorem~\ref{thm:energy_bal})
follows from the single identity $G\MPl^2 = \hbar c$, which is not an
empirical coincidence but an exact consequence of how the Planck units are
defined. Numerically:
\begin{equation}
  \frac{G\MPl^2}{\hbar c}
  = \frac{G \cdot (\hbar c / G)}{\hbar c}
  = 1.000\,000\,000\,000\,000
\end{equation}
to 15 significant figures (machine double precision).
The energy is not \emph{approximately} zero; it is \emph{algebraically}
zero from the definitions of $G$, $\hbar$, and $c$.

The physical significance: the universe contains $1.044 \times 10^{11}$~J
of matter energy per Planck cell at emergence, and an equal and opposite
$-1.044 \times 10^{11}$~J of gravitational energy.
The net is exactly zero. The apparent richness of the universe is a
rearrangement of nothing.

\subsection{Sensitivity to $\gstar$}

Although $E_\mathrm{cell}/E_\mathrm{thermal} = 15/\pi^2$ is independent
of $\gstar$, the absolute energy $E_\mathrm{cell} = \gstar\EPl/2$ is
linear in $\gstar$.
A change of one effective degree of freedom shifts the cell energy by
$\EPl/2 \approx 9.78 \times 10^{8}$~J, or about $0.94\%$ of the total.
For the SM value $\gstar = 106.75$,
\begin{equation}
  \frac{\partial E_\mathrm{cell}}{\partial \gstar} = \frac{\EPl}{2}
  = 9.780 \times 10^8\;\mathrm{J\;per\;effective\;DOF.}
\end{equation}
This makes $E_\mathrm{cell}$ a direct probe of the particle content at
emergence: any future measurement of the Planck-era energy density
constrains $\gstar$ without any other free parameter.

% ══════════════════════════════════════════════════════════════════════════════
\section{Quantitative Predictions}
\label{sec:predictions}
% ══════════════════════════════════════════════════════════════════════════════

Table~\ref{tab:predictions} assembles the quantitative predictions of TUO
that follow strictly from the two axioms plus SM particle content,
together with their observational status.

\begin{table}[ht]
\centering
\caption{TUO results from Axioms~\ref{ax:flat}--\ref{ax:zs} and SM content.
Status: $\checkmark\checkmark$ = TUO derives what standard cosmology assumes
as input; $\checkmark$ = quantitatively consistent with observation;
$\sim$ = qualitatively consistent, full calculation needed;
$\star$ = new distinguishing prediction; $\dagger$ = open problem.}
\label{tab:predictions}
\begin{tabular}{cp{3.2cm}p{4.2cm}p{3.0cm}c}
\toprule
\# & Result & TUO value & Observation & Status \\
\midrule
1 & Total energy $E_\mathrm{tot}$ & $= 0$ exactly (Theorem~\ref{thm:energy_bal}) &
  $\approx 0$ (flat universe) & $\checkmark\checkmark$ \\
2 & Spatial flatness $\Omega$ & $= 1$ exactly (Axiom~\ref{ax:flat}) &
  $1.0007\pm0.0037$~\cite{Planck2018} & $\checkmark\checkmark$ \\
3 & Equation of state $w$ & $= 1/3$ derived (Sec.~\ref{sec:expansion}) &
  $1/3$ (input to FRW) & $\checkmark\checkmark$ \\
4 & Hubble $H(\tPl)$ & $= 1/(2\tPl)$ derived (Sec.~\ref{sec:junction}) &
  $1/(2\tPl)$ (FRW input) & $\checkmark\checkmark$ \\
5 & Expansion law & $V \propto t^3$, $v < c$ always &
  $V \propto t^3$ (rad.\ dom.) & $\checkmark\checkmark + \star$ \\
6 & Energy per cell & $(\gstar/2)\EPl = 1.04\times10^{11}$\,J &
  $6.87\times10^{10}$\,J (thermal) & $\checkmark$ (factor $15/\pi^2$) \\
7 & Quantum correction & $\Delta V/V = 3(\lPl/ct)^2$ &
  Not yet measured & $\star$ \\
8 & $B-L = 0$ per generation & SM colour structure (Theorem~\ref{thm:bmL}) &
  SM anomaly cancellation & $\checkmark$ \\
9 & Proton stability & $\Delta(B-L)=0$ channels allowed &
  $\tau_p > 10^{34}$\,yr~\cite{SuperK2020} & $\star$ testable \\
10 & Baryon asymmetry & $B = L \neq 0$ permitted &
  $\eta = 6.12\times10^{-10}$ & $\sim$ $\dagger$ \\
11 & CMB uniformity & Simultaneous emergence (Sec.~\ref{sec:path_integral}) &
  $\Delta T/T \sim 10^{-5}$ & $\sim$ $\dagger$ \\
\bottomrule
\end{tabular}
\end{table}

% ══════════════════════════════════════════════════════════════════════════════
\section{Open Problems}
\label{sec:open}
% ══════════════════════════════════════════════════════════════════════════════

\begin{openq}[Baryon asymmetry]\label{oq:baryons}
TUO shows that $B - L = 0$ permits matter-only configurations.
It does not derive $\eta = 6.12\times10^{-10}$.
A full derivation requires QFT baryon-number dynamics at $\TPl$
with Sakharov-type washout factors.
\end{openq}

\begin{openq}[CMB power spectrum]\label{oq:cmb}
The saddle-point argument of Section~\ref{sec:path_integral} suggests
simultaneous emergence, giving $\Delta T/T = 0$ at leading order.
Within-cell quantum fluctuations give
$\Delta E_{\mathrm{cell}}/E_{\mathrm{cell}} \sim 1/\sqrt{\gstar} \approx 9.7\%$.
The observed $\Delta T/T \sim 10^{-5}$ requires $\sim 4$ orders of
magnitude of suppression through GR dynamics.
The specific calculation required is the two-point stress-energy
correlator $\langle T_{\mu\nu}(x)T_{\rho\sigma}(x')\rangle|_{\mathrm{ZS}}$
in a Planck-cell volume, propagated to horizon crossing.
\end{openq}

\begin{openq}[Energy density mismatch factor $15/\pi^2$]\label{oq:factor}
The TUO Heisenberg energy exceeds the FRW thermal energy by $15/\pi^2 \approx 1.52$.
Whether this is a physical effect (the pre-emergence state is colder than
equilibrium) or an artifact of the junction prescription is undetermined.
\end{openq}

\begin{openq}[Dark energy]\label{oq:lambda}
Zero-sum predicts $\Lambda_\mathrm{eff} = 0$ at leading order.
The residual $\Lambda_\mathrm{obs} \approx 10^{-122}\EPl^4$ is not derived
here and is left as an open problem for the companion QGD framework.
\end{openq}

\begin{openq}[SM particle content]\label{oq:sm}
TUO uses $\gstar = 106.75$ from experiment.
Whether SM content follows from the constrained path integral is open.
\end{openq}

\begin{openq}[why three dimensions]\label{oq:sm}
I have not as yet found the any meaningful reason behind the observed
three dimensions, so I cannot pose any hypothesis
\end{openq}


% ══════════════════════════════════════════════════════════════════════════════
\section{The Hot Big Bang Begins Here}
\label{sec:hbb}
% ══════════════════════════════════════════════════════════════════════════════

TUO is not a replacement for the Hot Big Bang (HBB) model.
It is a complement: TUO describes the pre-emergence phase and delivers
the HBB its initial conditions. This section states precisely where TUO
ends and the HBB begins, and verifies that the handoff is seamless.

\subsection{What the Hot Big Bang requires at $t = \tPl$}

The standard HBB model takes the following as empirical starting conditions
at $t = \tPl$~\cite{Weinberg2008,Kolb1990}:
\begin{enumerate}[leftmargin=2.5em,label=(\roman*)]
  \item Radiation equation of state: $w = 1/3$
  \item Hubble rate: $H(\tPl) = 1/(2\tPl)$ (radiation domination)
  \item Spatial flatness: $\Omega = 1$
  \item Full SM plasma: all $\gstar = 106.75$ effective degrees of freedom
  \item Thermal equilibrium at $T \sim \TPl$
\end{enumerate}
None of these are derived within the HBB itself.

\subsection{What TUO delivers}

\begin{theorem}[TUO-to-HBB handoff]\label{thm:handoff}
The maximum-fluctuation emergence configuration delivers all five HBB
initial conditions without additional assumptions.
\end{theorem}
\begin{proof}
\textit{(i) Equation of state.}
Wavepacket initial condition with $\Delta x = \lPl$ gives
$\Delta p = \MPl c/2$; relativity gives $v_0 = c$; therefore $w = 1/3$
(Section~\ref{sec:expansion}, derived not assumed).

\textit{(ii) Hubble rate.}
$H_\mathrm{TUO}(\tPl) = 1/(2\tPl)$ from the expansion law
(Eq.~\eqref{eq:H_junction}). In radiation-dominated FRW, $H = 1/(2t)$
exactly. Both are equal at $t = \tPl$.

\textit{(iii) Spatial flatness.}
Axiom~\ref{ax:flat} gives $k = 0$ exactly.

\textit{(iv) Full SM plasma.}
$E_\mathrm{cell} = \gstar\EPl/2 \approx 1.04 \times 10^{11}$~J exceeds
the pair-production threshold $2m_fc^2$ for every SM particle.
The heaviest (top quark, $m_t \approx 173\,\mathrm{GeV}$) has threshold
$5.46 \times 10^{-8}$~J; the ratio is $E_\mathrm{cell}/(2m_tc^2)
\approx 1.9 \times 10^{18}$.
All particle species are produced instantaneously.

\textit{(v) Thermal equilibrium.}
All $\gstar$ modes emerge simultaneously with the same Heisenberg energy
$\EPl/2$. This is, by definition, a thermal distribution at
$\TTUO = (15/\pi^2)^{1/4}\TPl \approx 1.11\,\TPl$.
No further thermalisation is required.
\end{proof}

\subsection{The junction}

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{ccc}
\textbf{TUO} ($t < \tPl$) & \textbf{at $t = \tPl$} & \textbf{Hot Big Bang} ($t > \tPl$) \\
\hline
Zero-sum pre-emergence & $\longleftrightarrow$ & Radiation domination \\
Wavepacket $\sigma(t)$, $v<c$ & & $a(t) \propto t^{1/2}$ \\
$w = 1/3$ (derived) & $w = 1/3$ & $w = 1/3$ (assumed) \\
$H = 1/(2\tPl)$ (derived) & $H$ continuous & $H = 1/(2t)$ \\
$k = 0$ (Axiom I) & $\Omega = 1$ & $\Omega = 1$ (assumed) \\
$\gstar = 106.75$ (all SM) & plasma established & HBB thermodynamics \\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{center}

\begin{remark}
The energy density mismatch by $15/\pi^2 \approx 1.52$ (Section~\ref{sec:stats})
does not break the continuity of $H$, $w$, or $k$, since $H$ is derived
from the expansion law independently of temperature.
As the initial maximum-fluctuation state thermalises, the effective
temperature relaxes to the value consistent with the Friedmann equation.
The resolution of this factor is Open Question~\ref{oq:factor}.
\end{remark}

% ══════════════════════════════════════════════════════════════════════════════
\section{Summary}
\label{sec:summary}
% ══════════════════════════════════════════════════════════════════════════════

The Theory of Universal Origins is a pre-Big-Bang framework based on two
axioms applied to the infinite-dimensional Fock space of Standard Model fields.
Its central object is the infinite zero matrix $\ZeroMat$; its central
equation is $\mathbf{Q}[\rhohat] = \ZeroMat$.

\noindent\textbf{Derived results} (from axioms alone):
\begin{enumerate}[leftmargin=2.5em]
  \item $w = 1/3$ --- the radiation equation of state
  \item $H(\tPl) = 1/(2\tPl)$ --- the Planck-era Hubble rate
  \item $\Omega = 1$ --- exact spatial flatness
  \item $V(t) \propto t^3$, $v < c$ always --- the expansion law
  \item $E_\mathrm{cell} = (\gstar/2)\EPl$, $E_\mathrm{grav} = -E_\mathrm{cell}$ --- zero total energy
  \item No-annihilation theorem --- matter-only stability
  \item $B-L = 0$, $Q = 0$ per SM generation --- anomaly identity
  \item $\TTUO = (15/\pi^2)^{1/4}\TPl$, independent of $\gstar$ --- universal temperature
  \item All five HBB initial conditions --- seamless handoff
\end{enumerate}

\noindent\textbf{Open problems} for future work:
baryon asymmetry $\eta$,
CMB power spectrum amplitude,
the $15/\pi^2$ energy-density factor,
dark energy,
SM particle content from axioms,
derivation of $N_\mathrm{gen} = 3$.

\section*{Acknowledgments}
The author thanks the Department of Physics, University of South Africa.
Numerical verification by Claude (Anthropic), February 2026.

% ══════════════════════════════════════════════════════════════════════════════
\begin{thebibliography}{99}
% ══════════════════════════════════════════════════════════════════════════════

\bibitem{Planck2018}
Planck Collaboration (N.~Aghanim et al.),
``Planck 2018 results VI. Cosmological parameters,''
\textit{Astron.\ Astrophys.}\ \textbf{641}, A6 (2020).
\href{https://doi.org/10.1051/0004-6361/201833910}{doi:10.1051/0004-6361/201833910}.

\bibitem{PlanckInflation2018}
Planck Collaboration (Y.~Akrami et al.),
``Planck 2018 results X. Constraints on inflation,''
\textit{Astron.\ Astrophys.}\ \textbf{641}, A10 (2020).
\href{https://doi.org/10.1051/0004-6361/201833887}{doi:10.1051/0004-6361/201833887}.

\bibitem{PDG2022}
R.~L.~Workman et al.\ (Particle Data Group),
``Review of Particle Physics,''
\textit{Prog.\ Theor.\ Exp.\ Phys.}\ \textbf{2022}, 083C01 (2022).
\href{https://doi.org/10.1093/ptep/ptac097}{doi:10.1093/ptep/ptac097}.

\bibitem{Guth1981}
A.~H.~Guth,
``Inflationary universe: A possible solution to the horizon and flatness problems,''
\textit{Phys.\ Rev.\ D}\ \textbf{23}, 347 (1981).
\href{https://doi.org/10.1103/PhysRevD.23.347}{doi:10.1103/PhysRevD.23.347}.

\bibitem{Linde1982}
A.~D.~Linde,
``A new inflationary universe scenario,''
\textit{Phys.\ Lett.\ B}\ \textbf{108}, 389 (1982).
\href{https://doi.org/10.1016/0370-2693(82)91219-9}{doi:10.1016/0370-2693(82)91219-9}.

\bibitem{SuperK2017}
Super-Kamiokande Collaboration (K.~Abe et al.),
``Search for proton decay via $p \to e^+ \pi^0$,''
\textit{Phys.\ Rev.\ D}\ \textbf{95}, 012004 (2017).
\href{https://doi.org/10.1103/PhysRevD.95.012004}{doi:10.1103/PhysRevD.95.012004}.

\bibitem{Weinberg2008}
S.~Weinberg,
\textit{Cosmology}
(Oxford University Press, Oxford, 2008).

\bibitem{Kolb1990}
E.~W.~Kolb and M.~S.~Turner,
\textit{The Early Universe}
(Addison-Wesley, Reading MA, 1990).

\bibitem{Lovelock1971}
D.~Lovelock,
``The Einstein tensor and its generalizations,''
\textit{J.\ Math.\ Phys.}\ \textbf{12}, 498 (1971).
\href{https://doi.org/10.1063/1.1665613}{doi:10.1063/1.1665613}.

\bibitem{Tryon1973}
E.~P.~Tryon,
``Is the Universe a Vacuum Fluctuation?''
\textit{Nature}\ \textbf{246}, 396 (1973).
\href{https://doi.org/10.1038/246396a0}{doi:10.1038/246396a0}.

\bibitem{Noether1918}
E.~Noether,
``Invariante Variationsprobleme,''
\textit{Nachr.\ Ges.\ Wiss.\ G\"ottingen}\ 235 (1918).
[\href{https://arxiv.org/abs/physics/0503066}{arXiv:physics/0503066}]

\bibitem{Matshaba2025}
R.~Matshaba,
``Universal Origins: The Zero-Sum Constraint,''
preprint, University of South Africa (2025).

\end{thebibliography}

\end{document}
